Precise Data Flow Analysis\\in the Presence of Correlated Method Calls

### 2015-03-09 sas2015/content/background.tex
ondrejs comments
move figures around to save half a page
proofreading pass

\subsubsection{Overview of the IFDS Algorithm}\label{sec:overviewifds}
The key idea behind the IFDS algorithm is that it is possible to represent any distributive
function $f$ from $2^D$ to $2^D$ by a \textit{representation relation} $R_f \subseteq 
(D \cup \{0\})
\times
(D \cup \{0\})$. The representation relation can be visualized as a bipartite graph
with edges from one instance of $D \cup \{0\}$ to another instance of $D \cup \{0\}$.
The IFDS algorithm uses such graphs to efficiently represent both the input data-flow functions
and the summary functions that it computes for called procedures.
Specifically, the representation relation $R_f$ of a function $f$ is defined as:
\[
R_f=\{(\mathbf0,\,\mathbf0)\}\cup
      \{(\mathbf0,\,d_j)\,|\,d_j\in f(\varnothing)\}\cup
      \{(d_i,\,d_j)\,|\,d_j\in f(\{d_i\}) \setminus f(\varnothing)\}.
\]
\begin{example}\label{ex:flowfun}
Given $D=\{u,\,v,\,w\}$ and $f(S) = S\setminus\{v\} \cup \{u\}$,
    the representation
 relation 
  $
    R_f=\{(\mathbf0,\,\mathbf0),\,(\mathbf0,\,u),\,(w,\,w)\}
$, which is depicted in Figure~\ref{fig:minigraph1}.
\end{example}

@@

@@

\subsubsection{Overview of the IDE Algorithm}\label{sec:ideoverview}
Just as any distributive function from $2^D$ to $2^D$ can be represented with a
representation relation, it is also possible to represent any distributive functions from
$\textsf{Env}(D,L)$
to
$\textsf{Env}(D,L)$
with a \textit{pointwise representation}. A pointwise representation is a bipartite graph
with the same nodes~\footnote{The IDE literature uses the symbol $\Lambda$ for the node that
    is denoted $\mathbf0$ in the IFDS literature. We use $\mathbf0$ throughout this paper for consistency.}
and edges as a representation relation, except that each edge is labelled
with a \textit{micro-function}, which is a function from $L$ to $L$.

@@

\begin{example}
    In the exploded supergraph in Figure~\ref{fig:cc_edgefn_example},
    the micro-functions are shown as labels on the graph edges.
    Every edge without an explicit label has the identity as its
    micro-function. The micro-functions on the three edges from
    the node \verb+return secret()+ to the node $\texttt{end}_\texttt{A.foo}$
    together represent the environment transformer 
    $\lambda e.e[\psi \to \lambda m.\bot \sqcap \lambda m.m]$.

>>>>

\subsubsection{Overview of the IFDS Algorithm}\label{sec:overviewifds}
The key idea behind the IFDS algorithm is that it is possible to represent any distributive
function $f$ from $2^D$ to $2^D$ by a \textit{representation relation} $R_f \subseteq 
(D \cup \{0\})
\times
(D \cup \{0\})$. The representation relation can be visualized as a bipartite graph
with edges from one instance of $D \cup \{0\}$ to another instance of $D \cup \{0\}$.
The IFDS algorithm uses such graphs to efficiently represent both the input data-flow functions
and the summary functions that it computes for called procedures.
Specifically, the representation relation $R_f$ of a function $f$ is defined as:
\[
R_f=\{(\mathbf0,\,\mathbf0)\}\cup
      \{(\mathbf0,\,d_j)\,|\,d_j\in f(\varnothing)\}\cup
      \{(d_i,\,d_j)\,|\,d_j\in f(\{d_i\}) \setminus f(\varnothing)\}.
\]
\vspace{-5mm}
\begin{example}\label{ex:flowfun}
Given $D=\{u,\,v,\,w\}$ and $f(S) = S\setminus\{v\} \cup \{u\}$,
    the representation
 relation 
  $
    R_f=\{(\mathbf0,\,\mathbf0),\,(\mathbf0,\,u),\,(w,\,w)\}
$, which is depicted in Figure~\ref{fig:minigraph1}.
\end{example}

@@

\input{figures/fig-cc_edge_fn_example}

@@

\subsubsection{Overview of the IDE Algorithm}\label{sec:ideoverview}
Just as any distributive function from $2^D$ to $2^D$ can be represented with a
representation relation, it is also possible to represent any distributive function from
$\textsf{Env}(D,L)$
to
$\textsf{Env}(D,L)$
with a \textit{pointwise representation}. A pointwise representation is a bipartite graph
with the same nodes~\footnote{The IDE literature uses the symbol $\Lambda$ for the node that
    is denoted $\mathbf0$ in the IFDS literature. We use $\mathbf0$ throughout this paper for consistency.}
and edges as a representation relation, except that each edge is labelled
with a \textit{micro-function}, which is a function from $L$ to $L$.

@@

\begin{example}
    In the exploded supergraph in Figure~\ref{fig:cc_edgefn_example},
    the micro-functions are shown as labels on the graph edges.
    Every edge without an explicit label has the identity as its
    micro-function. The micro-functions on the three edges from
    the node \verb+return secret()+ to the node $\texttt{end}_\texttt{A.foo}$
    together represent the environment transformer 
    $\lambda e.e[r \to \lambda m.\bot \sqcap \lambda m.m]$.

### 2015-03-09 sas2015/content/impl.tex
proofreading pass

%\paragraph{Lattice Elements}
In the correlated-calls transformation, the 
%Just like in the equivalence transformation $\transEq$, the exploded supergraph for $\transCC$ is the same as in the original IFDS problem. The 
lattice elements are maps from receivers to sets of types:
    $\lcc=\left\{\,m:\ R\to2^T\right\}$,
where $R$ is the set of receivers and $T$ is the set of all types.
For each receiver~$r$, the map gives an overapproximation of the possible
runtime types of $r$.
Sets of types are ordered by the superset relation, and this is lifted to maps from receivers
to sets of types, so the bottom element $\botcc$ maps every receiver to the set of all types,
and the top element $\topcc$ maps every receiver to the empty set of types.
During an actual execution, every receiver $r$ points to an object of some
runtime type. Therefore, a data-flow fact
is unreachable along any feasible path if its corresponding lattice element
maps any receiver to the empty set of types.

@@

\begin{example}
  Consider the program from Figure~\ref{list:ccexample}, whose exploded supergraph appeared in Figure~\ref{fig:cc_edgefn_example}.
%
  Returning a secret value in method \code{A.foo} creates a ``diagonal'' edge from the $\mathbf0$-fact to the method's return value $r$. 
  The diagonal edge is labeled with $\lambda m\,.\,\botcc$, so every receiver is mapped to the set of all types~$\bot_T$.
%  
On the end-return edge from \code{A.foo} to \code{main}, the set of types of \code{a} is restricted by the micro function $\lambda m\,.\,m[\code a\to m(\code a)\cap\{\code A\}]$ corresponding to the assignment of the return value $r$ to \code v.
%
%Similarly,
On the call-start edge from \code{main} to \code{B.bar}, the possible types
of \code{a} are further restricted by the micro-function
$\lambda m\,.\,m[\code a\to m(\code a)\cap\{\code B\}]$.
on the edge that passes the argument \code{v} to the parameter \code{s}.
The composition of these micro functions results in the empty set as the possible
types \code{a}, indicating that this
%data-flow
path 
%that would
%result in an information leak 
is 
%actually
infeasible.
\end{example}

@@

In the equivalence transformation lattice $L^\equiv$, $\bot$ means reachable
and $\top$ means unreachable.
Therefore, an IDE solution $\rho$ computed using $\transEq$ is converted to
and IFDS solution as:
  $\backEq(\rho) = \lambda n.\{d\mid\rho(n)(d)\ne\top\}$.
%a result $\rho=\resultIDE(\transEq(P))$ of the equivalence IDE analysis
%is converted to an IFDS result as follows:
%\begin{equation}\label{eq:transEqDef}
  %\backEq(\rho) = \lambda n.\{d\mid\rho(n)(d)\ne\top\}.
%\end{equation}
%
%\otodo{I don't understand this comment. What does it mean?}
%
In the correlated-calls transformation lattice $\lcc$, a map that maps any receiver
to the empty set of possible types means that the corresponding data-flow path is
infeasible. Therefore, an IDE solution $\rho$ computed using
$\transCC_S$ is converted to an IFDS solution as:
$\backCC(\rho) = \lambda n.\{d\mid\forall r\in S\,.\,\rho(n)(d)(r)\neq\top_T\}$.

@@

%\otodo{Give the list of criteria explicitly in the proof in the appendix.}
%\begin{enumerate}
	%\item There is a representation for the identity and top functions.
  %\item The representation is closed under the meet and composition operations.
  %\item The micro functions form a finite-height lattice.
  %\item The apply, meet, composition, and equality-check operations can be computed in constant time (independent of $E$ and $D$).
  %\item There is a constant bound on the storage space for a micro function representation.
%\end{enumerate}
%We prove these conditions with the following lemma:
\begin{restatable}{lemma}{efficient}\label{lem:efficient}
  The correlated-call representation of a micro function is efficient and the required operations
  on micro-functions can be computed in time $O(\rcc T)$.
\end{restatable}

@@

\subsection{Implementation of the Correlated-Calls Analysis}
We implemented the correlated-calls analysis in Scala~\cite{odersky2004overview}. 
Our implementation analyzed Java bytecode from input programs written
in Java.
%The input programs that we analyzed were written in Java. Our implementation analyzed the
%corresponding Java bytecode.
We used WALA~\cite{fink2012wala}
to retrieve information about an input program, such as its control-flow supergraph and the set of receivers and their types.
Since WALA does contain an implementation of the IDE algorithm,
we implemented it from scratch.
%currently only contains an implementation of IFDS, we implemented IDE from scratch.
We will submit our implementation to the artifact evaluation committee, and we are working on contributing it officially to WALA. % Instead of using WALA's IFDS implementation, to run an IFDS problem, we converted it to an IDE problem and used our own IDE solver.

>>>>

%\paragraph{Lattice Elements}
In the correlated-calls transformation, the 
%Just like in the equivalence transformation $\transEq$, the exploded supergraph for $\transCC$ is the same as in the original IFDS problem. The 
lattice elements are maps from receivers to sets of types:
    $\lcc=\left\{\,m:\ R\to2^T\right\}$,
where $R$ is the set of receivers and $T$ is the set of all types.
For each receiver~$r$, the map gives an overapproximation of the possible
runtime types of $r$.
Sets of types are ordered by the superset relation, and this is lifted to maps from receivers
to sets of types, so the bottom element $\botcc$ maps every receiver to the set of all types,
and the top element $\topcc$ maps every receiver to the empty set of types.
During an actual execution, every receiver $r$ points to an object of some
runtime type. Therefore, a data-flow fact
is unreachable along a given path if its corresponding lattice element
maps any receiver to the empty set of types.

@@

\begin{example}
  Consider the program from Figure~\ref{list:ccexample}, whose exploded supergraph appeared in Figure~\ref{fig:cc_edgefn_example}.
%
  Returning a secret value in method \code{A.foo} creates a ``diagonal'' edge from the $\mathbf0$-fact to the method's return value $r$. 
  The diagonal edge is labeled with $\lambda m\,.\,\botcc$, so every receiver is mapped to the set of all types~$\bot_T$.
%  
On the end-return edge from \code{A.foo} to \code{main}, the set of types of \code{a} is restricted by the micro function $\lambda m\,.\,m[\code a\to m(\code a)\cap\{\code A\}]$ corresponding to the assignment of the return value $r$ to \code v.
%
%Similarly,
On the call-start edge from \code{main} to \code{B.bar}, the possible types
of \code{a} are further restricted by the micro-function
$\lambda m\,.\,m[\code a\to m(\code a)\cap\{\code B\}]$
on the edge that passes the argument \code{v} to the parameter \code{s}.
The composition of these micro functions results in the empty set as the possible
types of \code{a}, indicating that this
%data-flow
path 
%that would
%result in an information leak 
is 
%actually
infeasible.
\end{example}

@@

In the equivalence transformation lattice $L^\equiv$, $\bot$ means reachable
and $\top$ means unreachable.
Therefore, an IDE solution $\rho$ computed using $\transEq$ is converted to
an IFDS solution as:
  $\backEq(\rho) = \lambda n.\{d\mid\rho(n)(d)\ne\top\}$.
%a result $\rho=\resultIDE(\transEq(P))$ of the equivalence IDE analysis
%is converted to an IFDS result as follows:
%\begin{equation}\label{eq:transEqDef}
  %\backEq(\rho) = \lambda n.\{d\mid\rho(n)(d)\ne\top\}.
%\end{equation}
%
%\otodo{I don't understand this comment. What does it mean?}
%
In the correlated-calls transformation lattice $\lcc$, a map that maps any receiver
to the empty set of possible types means that the corresponding data-flow path is
infeasible. Therefore, an IDE solution $\rho$ computed using
$\transCC_S$ is converted to an IFDS solution as:
$\backCC(\rho) = \lambda n.\{d\mid\forall r\in S\,.\,\rho(n)(d)(r)\neq\top_T\}$.

@@

%\otodo{Give the list of criteria explicitly in the proof in the appendix.}
%\begin{enumerate}
	%\item There is a representation for the identity and top functions.
  %\item The representation is closed under the meet and composition operations.
  %\item The micro functions form a finite-height lattice.
  %\item The apply, meet, composition, and equality-check operations can be computed in constant time (independent of $E$ and $D$).
  %\item There is a constant bound on the storage space for a micro function representation.
%\end{enumerate}
%We prove these conditions with the following lemma:
\begin{restatable}{lemma}{efficient}\label{lem:efficient}
  The correlated-call representation of a micro function is efficient according to the IDE criteria and the required operations
  on micro-functions can be computed in time $O(\rcc T)$.
\end{restatable}

@@

\subsection{Implementation of the Correlated-Calls Analysis}
We implemented the correlated-calls analysis in Scala~\cite{odersky2004overview}. 
Our implementation analyzed Java bytecode compiled from input programs written
in Java.
%The input programs that we analyzed were written in Java. Our implementation analyzed the
%corresponding Java bytecode.
We used WALA~\cite{fink2012wala}
to retrieve information about an input program, such as its control-flow supergraph and the set of receivers and their types.
Since WALA does not contain an implementation of the IDE algorithm,
we implemented it from scratch.
%currently only contains an implementation of IFDS, we implemented IDE from scratch.
We will submit our implementation to the artifact evaluation committee, and we are working on integrating it and contributing it officially to WALA. % Instead of using WALA's IFDS implementation, to run an IFDS problem, we converted it to an IDE problem and used our own IDE solver.
